---
title: "Assignment 1"
author: "Sherry Mu"
output:
  word_document: default
  html_document: default
---

#Question 1
Assume Bill's annual repayments are $R_{B}$ and the interest rate is i. 

Since the first instalment is due 19/03/2020, know that R is paid annually in arrear and the last repayment will be due 19/03/2050.

####Firstly, set the function to calculate annuities
```{r}
Annuity <- function(i,n){
  (1-(1+i)^(-n))/i
}
```

####Thus, calculate $R_{B}$ by sustituting everyting in the equation $$600000 = R_{B}a_{30}^{@i}$$
```{r}
i <- 0.035
n <- 30
Rb <- 600000/Annuity(i, n)
```

####Bill's annual repayments are
```{r, echo=FALSE}
Rb
```

#Question 2
##(a)
Let Scott's loan repayments be $R_{S}$

####$R_{S}$ can thus be calculated by $$600000-100000(1+i)^{-n} = (R_{S}+100000i)a_{30}^{@i}$$
```{r}
i <- 0.035
n <- 30
Rs <- (600000-100000*(1+i)^(-n))/Annuity(i, n) - 100000*i
```

####Scot's annual repayments are
```{r, echo=FALSE}
Rs
```

##(b)
If Scot paid $27185.67 together with offset interest of $3500 per annual for 30 years, the PV of all his instalments will be $$PV_{S} = (R_{S}+1000000i)a_{30}^{@i}$$
```{r}
PV <- (Rs + 100000*i)*Annuity(i,n)
PV
```
which is greater than his net balance, 500000. This indicates that Scot can pay his repayments for less than 30 years to cover his loan.

Now, want to find the exat years that the loan extinguished
```{r}
fn <- function(n){
  (600000-100000) - (Rs+100000*i)*Annuity(i,n)
}
N <- uniroot(fn, c(0,30))
N$root
```

This means that Scot will pay $30685.67 annually since 19/03/2020 for 24 years, and at the 25th year, on 19/03/2045, he will need to pay
```{r}
fn(24)*(1+i)^25
```

#Question 3
##(a)
As Bill will pay Theresa $1000 annually for 30 years, the total PV of her commissions will be
```{r}
i <- 0.035
n <- 30
PVt <- 1000*Annuity(i,n)
PVt
```

##(b)
The effective interest rate to Bill will be lower than 3.5%.

This is because combining with Theresa's total commission payments, the PV of Bill's total loan will be increased. However, Bill is still paying $R_{B}$ without adding the $1000 commission payment each year. This means that the $1000 will be paid by sacrificing part of his inerest earnings, and hence the effective interest that the bank requires on its fund must be lower than 3.5%.

##(c)
####Firstly, assume the effective rate to Bill is j and we want to solve $$600000 + 1000a_{30}^{@i} = R_{B}a_{30}^{@j}$$
```{r}
n <- 30
fi <- function(j){
  600000+PVt - Rb*Annuity(j,n)
}
```

####By setting thr function $f = 600000 + 1000a_{30}^{@i} - R_{B}a_{30}^{@j}$ equals 0, we can hence find the effective rate j
```{r}
j <- uniroot(fi, interval = c(0.0000001, 0.035))
```

Theoretically the range of j will be between 0 and 3.5%, but in Question 1 the "Annuity" function is set divided i. Hence in the uniroot function above, we just try to set the lower end point of j extremely small to make the value valid.

####The effective rate is
```{r, echo=FALSE}
j$root
```

##(d)
As in (a), 3.5% is indicated a valuation interest rate, Bill's loan will be calculated as 600000 together with an annuity of 1000 p.a. for 30 years at j (the effective rate calculated above).

Bill's new annual payment will be
```{r}
n <- 30
Rb_n <- (600000+1000*Annuity(j$root,n))/Annuity(j$root,n)
Rb_n
```

Hence, his annual payment is increased by
```{r}
Rb_n - Rb
```

#Question 4
####Firstly, set up a function to calculate age in years
```{r}
age <- function(from, to){
  n <- as.POSIXlt(c(from, to))
  age <- n$year[2]-n$year[1]
  age
}
```

##Life A
###(a) age next birthday at death
```{r}
age("1965/05/16", "1995/05/16")
```

###(b) age last birthday on 1 January before death
```{r}
age("1965/05/16", "1993/05/16")
```

###(c) age at birthday in the policy year of death
```{r}
age("1965/05/16", "1994/05/16")
```

##Life B
###(a) age next birthday at death
```{r}
age("1922/03/01", "1994/03/01")
```

###(b) age last birthday on 1 January before death
```{r}
age("1922/03/01", "1992/03/01")
```

###(c) age at birthday in the policy year of death
```{r}
age("1922/03/01", "1993/03/01")
```

##Life C
###(a) age next birthday at death
```{r}
age("1942/09/21", "1996/09/21")
```

###(b) age last birthday on 1 January before death
```{r}
age("1942/09/21", "1995/09/21")
```

###(c) age at birthday in the policy year of death
```{r}
age("1942/09/21", "1996/09/21")
```
